Draft Documents

### 2022-11-17 src/draft.Rmd
update
Update draft.Rmd

---
title: "Draft Documents"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: citations.bib
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
# inline hook to 2 digits
inline_hook <- function(x) {
    if(is.numeric(x)) x <- round(x, 2)
    paste(as.character(x), collapse=", ")
}
knit_hooks$set(inline = inline_hook)
# code chunk out put to 
options(digits = 5)
```

```{r, message = F}
library(needs)

needs("dplyr", "MASS", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet", "ggforce", "lmtest")

@@

# 1. Introduction (To be done)

@@

## 2.3. Descriptive Analytics
Distribution of the predicted variable `pov`
```{r}
ggplot(countries, aes(x = pov)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density(alpha = .2, fill = "blue")
```
The graph displays a decreasing rate as poverty indicator increasing. This might not be representative of the current state of poverty in the world, but of the number presented in our data. For example, more recent data is likely to be more inclusive than ancient data, when poverty is more prevalent. We should look at data from the same period.

```{r}
# number of data point available from 1967 to 2021
ggplot(poverty.headcount, aes(x = year)) +
  geom_histogram(bins = 30)

@@

ggplot(pov.98.18, aes(x = pov)) +
  geom_histogram(aes(y = ..density..), bins = 30) +
  geom_density(alpha = .2, fill = "blue") +
  facet_grid(cols = vars(year))
```

@@

# World
ggplot(pov.reg %>% filter(country.code == "WLD"),
       aes(x = year, y = pov)) +
  geom_point() +
  geom_smooth(formula = y~x, method = loess)
```

@@

```{r}
ggplot(pov.reg %>% filter(country.code != "WLD"),
       aes(x = year, y = pov, color = country.name)) +
  geom_point() +
  geom_smooth(formula = y~x, method = loess) +
  facet_grid(rows = vars(type))
```

@@

## 2.4. Assumption Check
We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data. [Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
**1. Linear relationship:**  
Use correlation matrix to check linearity
```{r, warning = FALSE}
# select only numeric data
countries.num <- countries %>% select(where(is.numeric))
chart.Correlation(countries.num, histogram = TRUE, pch = 19)
```
Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
```{r}
# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
countries.num2 <- countries.num %>% 
  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
  filter(!is.na(value) & !is.na(pov))

@@

# for (indvar in ivs) {
#   print(
#     ggplot(countries.num2 %>% 
#            filter(variable == indvar),
#          aes(x = value, y = pov)) +
#     geom_point() +
#     geom_smooth(formula = y~x, method = lm) +
#     labs(title = paste("Relationship pov ~", indvar),
#          x = indvar,
#          y = "pov")
#   )
# }

p <- ggplot(countries.num2, 
     aes(x = value, y = pov)) +
geom_point() +
geom_smooth(formula = y~x, method = lm) +
labs(title = paste("pov ~ variable"),
     x = "variable",
     y = "pov") +
facet_wrap_paginate(~ variable,
                    ncol = 3,
                    nrow = 1,
                    scales = "free")

requiredPages <- n_pages(p)

for (i in 1:requiredPages) {
  ggplot(countries.num2,
         aes(x = value, y = pov)) +
  geom_point() +
  geom_smooth(formula = y~x, method = lm) +
  labs(title = paste("pov ~ variable"),
       x = "variable",
       y = "pov") +
  facet_wrap_paginate(~ variable,
                      ncol = 3,
                      nrow = 1,
                      scales = "free",
                      page = i) -> p
  print(p)
}
```

@@

for (indvar in logIvs) {
  print(
    ggplot(countries.num2 %>% 
           filter(variable == indvar),
         aes(x = log(value), y = pov)) +
    geom_point() +
    geom_smooth(method = lm) +
    labs(title = paste0("pov ~ log(", indvar, ")"),
         x = paste0("log(", indvar, ")"),
         y = "pov")
  )
}
```

@@

ggplot(gdr.eql,
       aes(x = value, y = pov)) +
  geom_boxplot() +
  labs(title = paste("pov ~ gdr.eql"),
       x = "gdr.eql",
       y = "pov")
```
There's a significant correlation between gender equality and poverty.  
Correlation coefficients between `pov` and predictors.

@@

## 2.5. Data Source

# 3. Model Selection and Interpretation

We can now conduct linear regressions following the approaches mentioned above to address missing values issues.

## 3.1. Use Complete Cases (To be done)
###  3.1.1. Model Fitting
###  3.1.2. Assessment
###  3.1.3. Interpretation

## 3.2. Selectively remove variables with high missing rate (To be done)
**1. About selectively removing variables with high missing rate**

Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables. This section attempts to build a model on variables that have missing rates of below 35%.

The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models. Hence, we transformed the data with the mutate() function before building the machine learning model.
```{r}
subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))

subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))

subset1 <- na.omit(subset1)
```
We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step.

```{r}
#helper function
sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

@@

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}

df.countries0 <- standardise(subset1)
df.countries <- df.countries0
head(df.countries)
```

###  3.2.1. Model Fitting
##### A. Ordinary Linear Regression
First, we split the data into training and testing data. The training dataset contains 80% of the initial dataset, and the testing data contains 20% of the whole dataset. 
```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

```
Our first model includes all the variables with missing rate lower than 35%. This model is assigned to `lm_1`.
```{r}
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05. We can imply that these variables do not provide a high significance to our model and are thus able to be removed. To avoid removing significant variables, we shall conduct the removal of the variables deliberately. In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|). Hence, we remove `lgdp.dflt`.
```{r}
lm_2 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
summary(lm_2)
```
Removing `lfdi`,
```{r}
lm_3 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
summary(lm_3)
```
Further removing `hlth`,
```{r}
lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
summary(lm_4)
plot(lm_4, which = 5)
```
Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.

##### B. VIF

##### c. Regularization
Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model. 
Later on, we will compare the results of all the models we generated.
```{r}
data0 <- subset1
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y <- data.matrix(data$pov)

```
Dividing into training and testing datasets.
```{r}
set.seed(123)
rate <- 0.6
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]
```
Further dividing into `train.x`, `train.y`, `test.x`, and `test.y`.
```{r}
training <- na.omit(training)
train.x <- model.matrix(pov ~., data = training)[,-1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```
Using the lambda value that yields the lowest MSE value. 
```{r}
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)

t(coef(glm_ridge))
```

```{r}
#train the ridge regression model using lambda.min

glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
t(coef(glm_Ridge))
```

Function to evaluate the results 
```{r}
eval_results <- function(fit, true) {
  actual <- data.matrix(true)
  SSE <- sum((actual - fit)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE/SST
  data.frame(
    MSE = MSE(fit, true),
    MAE = MAE(fit, true),
    RMSE = RMSE(fit, true),
    MAPE = MAPE(fit, true),
    R2  = R_square

  )
}
```

```{r}
fit <- predict(glm_Ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)

@@

```{r}
fit <- predict(glm_Ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)

@@

```{r}
summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
knitr::kable(summary_Ridge, digits = 3)

###  3.2.2. Assessment
###  3.2.3. Interpretation

## 3.3. Update the data set as we select variables (To be done)
###  3.3.1. Model Fitting
###  3.3.2. Assessment
###  3.3.3. Interpretation

## 3.4. Imputation
**1. What is imputation**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), and use other variables to predict the missing values (much like regression). Since we observed some correlations during descriptive analysis, we expect the imputation algorithm to work well.

We use **Multivariate Imputation By Chained Equations (MICE)** as primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks). In order to use this technique, we have to assume that the "missingness" of a field can be explained by the values in other columns, (e.g., If the countries is in North America, it's more likely to be missing as we have discovered in 2.2). The general ideas of the algorithm is to fill in the missing values, and improve it iteratively until predicted values converse to a stable point. The algorithmic details of MICE is very concisely (and enthrallingly) explained in @micealgo.

We use relatively small parameters in the interest of time.

@@

###  3.4.1. Model Fitting
##### A. Ordinary Linear Regression
We generated 3 sets of imputated (training) data. We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). With all the generated data sets.

```{r}
formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'

# Using single imputed data set
fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))

# Build models with all generated data set and pool the estimates
fit2 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
fit2.combined <- pool(fit2)
# dummy lm model
fit2.dummy <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
# replace coefficients of dummy model
name.coef <- names(fit2.dummy$coefficients)
fit2.dummy$coefficients <- fit2.combined$pooled$estimate
names(fit2.dummy$coefficients) <- name.coef
```
Helper function to calculate R-Squared
```{r}
r2 <- function(pred, orig) {
   RSS <- sum((pred - orig)^2)
   TSS <- sum((orig - mean(orig))^2)
   R2 <- 1 - RSS/TSS
   return (R2)
}

adjR2 <- function(pred, orig, k) {
  R2 <- r2(pred, orig)
  n <- length(pred)
  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
  return (adjr2)
}
```
Adjusted R-Squared on train and test
```{r}
fit1.summ <- lapply(fit1, \(f) {
  # number of variables
  k <- length(f$coefficients) - 1
  # predict value of test
  pred <- predict(f, countries.test.4)
  test.resid <- pred - countries.test.4$pov
  data.frame(
    `Train MSE` = mean(summary(f)$residuals^2),
    `Test MSE` = mean(test.resid^2),
    `Train MAE` = mean(abs(summary(f)$residuals)),
    `Test MAE` = mean(abs(test.resid)),
    `Train Adj.R2` = summary(f)$adj.r.squared,
    `Test Adj.R2` = adjR2(pred, countries.test.4$pov, k)
  )
})

# number of variables
k <- length(fit2.dummy$coefficients) - 1
# number of variables
fit2.pred <- predict(fit2.dummy, countries.test.4)
test.resid <- fit2.pred - countries.test.4$pov
fit2.summ <- data.frame(
  `Train MSE` = mean(summary(fit2.dummy)$residuals^2),
  `Test MSE` = mean(test.resid^2),
  `Train MAE` = mean(abs(summary(fit2.dummy)$residuals)),
  `Test MAE` = mean(abs(test.resid)),
  `Train Adj.R2` = pool.r.squared(fit2, adjusted = T)[,'est'],
  `Test Adj.R2` = adjR2(fit2.pred, countries.test.4$pov, k)
)

res <- do.call(rbind.data.frame, fit1.summ)
res <- rbind(res, fit2.summ)
res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)

kable(res)
```

Pooled model slightly improve performance. Test data fitting has better MSE and MAE than train data. However, Adjusted R-Squared is lower in test data. There might be some over-fitting in our models.

```{r}
# find coef of variable other than country.code
var <- c("year","incomeL","incomeLM","incomeUM","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","pop.gwth.urban","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")

res <- do.call(rbind, lapply(fit1, \(f) f$coefficients[var]))
res <- rbind(res, fit2.dummy$coefficients[var])
res <- cbind(data.frame(Set = c(1:3, "Pooled")), res)
kable(res)
```

In contradiction to expectation, `hlth` is positively correlated with `pov`, and lower-middle income countries (incomeLM) are less poor than high income countries (incomeH). `pop.gwth.total` is negatively related to `pov`, while `pop.gwth.rural` and `pop.gwth.urban` are positively related. They might be indications of over-fitting model.

```{r}
# fit model on imputated set no. 1
summary(fit1[[1]])
```

There are also some weak variable, we should perform some variable selection.

##### B. VIF

Our data set contains a lots of variables. We can perform some variable selection to reduce over-fitting.

```{r}
gvif <- lapply(fit1, vif)
gvif
```

[Some forum](https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif) suggested that we should use the standard $GVIF^{1/(2 \cdot df)} < 2$ as equivalent to $GVIF < 4$ to account for high degree of freedom of some variables.

```{r}
# adjusted gvif higher than 3
lapply(gvif, \(g) {
  g[g[,3] > 3, 3]
})
```

Remove the highest-ranked variables (`pop.gwth.total`, `pop.gwth.urban`, `pop.gwth.rural`, `gdp.pc`, `lgdp.pc`) and rebuild the models.

```{r}
formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+trade+lfdi+lgdp.dflt'

# Using single imputed data set
fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))
```

Re-run GVIF analysis

```{r}
gvif <- lapply(fit1, vif)

lapply(gvif, \(g) {
  g[g[,3] > 2, 3]
})
```

There're still some variable with adjusted GVIF > 2. We are interested in the effects of `hlth` (expenditure in Healthcare), and `mil` (expenditure in Military), so we won't remove those variables. We can remove `trade`.

```{r}
# remove trade
formula.str <- 'pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+gdp.dflt+gcf+lfdi+lgdp.dflt'

# Using single imputed data set
fit1 <- lapply(1:countries1.imputed$m, \(i) lm(as.formula(formula.str), data = complete(countries1.imputed, i)))

# gvif
gvif <- lapply(fit1, vif)

>>>>

---
title: "Draft Documents"
author: "Team 1 (Hedgehog, Callista, Imtiyaaz, Issac)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
bibliography: citations.bib
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)
```

```{r, message = F}
# Package names
# packages <- c("MASS", "knitr", "readr", "tidyr", "dplyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "needs")
# 
# # Install packages not yet installed
# installed_packages <- packages %in% rownames(installed.packages())
# if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }
# 
# # Packages loading
# invisible(lapply(packages, library, character.only = TRUE))
# 
# prioritize(dplyr)

library(needs)

needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")

@@

# 1. Introduction (To be done)
The project aim is to explore different models using the same dataset to evaluate which method will be the best to build a model for the dependent variable. The dataset we are using for this project has 1901 data values from 23 different variables. Our plan for this project is to first build a model with all the variables in the dataset. We will then follow it up by builidng a model without the missing values and build a ridge regression and Lasso regression model. Our second approach will be to selectively remove variables from the model and to build the best model  using regularisation and regression. Our next approach was to add the variables into the model one at a time to see if a different model can be achieved. Lastly, we performed imputation on our model to tackle the missing values present in our dataset.We then compared all of this approaches to come to a conclusion.

@@

## 2.3. Descriptive Analytics
Distribution of the predicted variable `pov`
```{r}
ggplot(countries, aes(x = pov)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density(alpha = .2, fill = "blue")
```
The graph displays a decreasing rate as poverty indicator increasing. This might not be representative of the current state of poverty in the world, but of the number presented in our data. For example, more recent data is likely to be more inclusive than ancient data, when poverty is more prevalent. We should look at data from the same period.

```{r}
# number of data point available from 1967 to 2021
ggplot(poverty.headcount, aes(x = year)) +
  geom_histogram()

@@

ggplot(pov.98.18, aes(x = pov)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density(alpha = .2, fill = "blue") +
  facet_grid(cols = vars(year))
```

@@

# World
ggplot(pov.reg %>% filter(country.code == "WLD"),
       aes(x = year, y = pov)) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
ggplot(pov.reg %>% filter(country.code != "WLD"),
       aes(x = year, y = pov, color = country.name)) +
  geom_point() +
  geom_smooth() +
  facet_grid(rows = vars(type))
```

@@

## 2.4. Data Source

# 3. Model Selection and Interpretation
## 3.1. Assumption Check
We can conduct some preliminary checks on linearity and correlation of predictors to have a better picture of the data. [Checklist](https://godatadrive.com/blog/basic-guide-to-test-assumptions-of-linear-regression-in-r)
**1. Linear relationship:**  
Use correlation matrix to check linearity
```{r, warning = FALSE}
# select only numeric data
countries.num <- countries %>% select(where(is.numeric))
chart.Correlation(countries.num, histogram = TRUE, pch = 19)
```
Which is utterly intelligible, but a majority of the fitted lines are linear at first glance. We should render separate graphs for the relationship between `pov` & other variables.
```{r}
# lengthen data table to variable-value pairs, with pov as predicted variable and others as predictive 
countries.num2 <- countries.num %>% 
  pivot_longer(cols = 3:ncol(.), names_to = "variable", values_to = "value") %>% 
  filter(!is.na(value) & !is.na(pov))

@@

for (indvar in ivs) {
  print(
    ggplot(countries.num2 %>% 
           filter(variable == indvar),
         aes(x = value, y = pov)) +
    geom_point() +
    geom_smooth(method = lm) +
    labs(title = paste("Relationship pov ~", indvar),
         x = indvar,
         y = "pov")
  )
}
```

@@

for (indvar in logIvs) {
  print(
    ggplot(countries.num2 %>% 
           filter(variable == indvar),
         aes(x = log(value), y = pov)) +
    geom_point() +
    geom_smooth(method = lm) +
    labs(title = paste0("Relationship pov ~ log(", indvar, ")"),
         x = paste0("log(", indvar, ")"),
         y = "pov")
  )
}
```

@@

ggplot(gdr.eql,
       aes(x = value, y = pov)) +
  geom_boxplot() +
  geom_smooth(method = lm) +
  labs(title = paste("Relationship pov ~ gdr.eql"),
       x = "gdr.eql",
       y = "pov")
```
There's a significant correlation between gender equality and poverty.  
Correlation coefficients between `pov` and predictors.

@@

## 3.2. Ordinary Multiple Linear Regression

We conduct a normal linear regression, following the approaches mentioned above to address missing values issues.

### 3.2.1. Use Complete Cases (To be done)
#### 3.2.1.1. Model Fitting
#### 3.2.1.2. Assessment
#### 3.2.1.3. Interpretation

```{r}
library(tidyverse) # for data manipulations and visualisations
library(ggpubr) # for correlation coefficient on plot
library(gridExtra) # for arranging plots 
library(e1071) # for skewness
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
library(corrplot)
library(car)
library(DescTools)
library(glmnet)
```

First we will write the function "eval_results(true,predict) to evaluate the models
```{r}
eval_results <- function(true, predict) {
  actual <- data.matrix(true)
  SSE <- sum((predict - actual)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE / SST
  data.frame(
    MSE = MSE(predict, true),
    MAE = MAE(predict, true),
    RMSE = RMSE(predict, true),
    Rsquare = R_square
  )
}
```
First we read the csv file for the countries containing the dataset. We will then put the variables we need in to subset1.
```{r}
df <-read.csv('../data/countries1.csv')
df
subset1 <- df %>% select("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc")
subset1
```

Next we will build a SLR model. We will put 80% of dataset to train and the rest to test.
```{r}
set.seed(10)
index <- sort(sample(x = nrow(subset1), size = nrow(subset1)*.8))
train <- df[index, ]
test <- df[-index, ]
```

The model output will be placed in lm.slr and can be viewed in summary(lm.slr). THe r-squared value is 0.9619.
```{r}
lm.slr <- lm(pov ~ country.code+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc,data = train)
summary(lm.slr)

```
Next we will be performing standardisation.However not all the variables are numerical. Therefore to overcome it we build the function below. This will allow us to perform standardisation to the dataset.
```{r}
subset1$income <- factor(subset1$income)
subset1$country.code <- factor(subset1$country.code)

sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

@@

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}
```

Next we will perform standardistation to the dataset.
```{r}
str(subset1)
df0 <- subset1
subset1
apply(subset1, 2, sd)

@@

Next we will plot a pov vs income scatter plot to see the effect of income on pov. From the scatter plot, we can deduce that the low income countries have a higher pov while the higher income countries have low pov. This is consistent with our intuition that countries that are richer do not have as much poverty issue as the poorer countries.
```{r}
ggplot(data = subset1, aes(x = income, y = pov)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Pov vs. Income ",
       x = "income",
       y = "pov")

```
Next we will build a MLR model with data from subset 1 to see the relationship between dependent variable:pov and the other variables

```{r}
model1 <- lm(pov ~ ., data = subset1)
summary(model1)
vif(model1)

```
Next we will perform regularisation. We will split the data to predictor variable "train.x" and response variables "train.y"
```{r}
subset2 <- subset(subset1,select =c("pov","country.code","year","income","edu.total","hlth","mil","fdi","lbr.part","unemp","pop.gwth.total","pop.gwth.rural","gdp.dflt","gcf","trade","gdp.pc","lfdi","lgdp.dflt","lgdp.pc"))
subset2 <- na.omit(subset2)
subset2
train.x <- data.matrix(subset(subset2, select = -pov))
train.y = data.matrix(subset2$pov)
```

Next we will perform splitting of data to train and test.
```{r}
split_train_test <- function(data) {
set.seed(1984)
# split data
seeds <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() == 1)

plants <- data %>% 
  group_by(country.code) %>% 
  filter(row_number() != 1)

isComplete <- which(complete.cases(plants))
idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))

training <- plants[-idx,] %>% 
  rbind(seeds)
test <- plants[idx,]

return(list(training, test))
}

training <- split_train_test(subset2)[[1]]
test <- split_train_test(subset2)[[2]]

ntrain <- 1:nrow(training)
combine <- rbind(training, test)
combine.mtrx <- model.matrix(pov ~ ., data = combine)
train.mtrx <- combine.mtrx[ntrain,]
test.mtrx <- combine.mtrx[-ntrain,]

train.x <- train.mtrx[,-1]
train.y <- training$pov

test.x <- test.mtrx[,-1]
text.y <- test$pov
```

We then build a MLR model with the training model to train it.

```{r}
model3 <- lm(pov ~., data = training)
summary(model3)

@@

```{r}
fit <- model3$fitted.values
true <- training$pov
summary_MLR3_train <- eval_results(true, fit)
```

We then evaluate the model between train and test
```{r}
fit <- predict(model3, newdata=test)
true <- test$pov
summary_MLR3_test <- eval_results(true, fit)
```

We can see the evaluation of the model from the function below. From the train model, the R^squared value will be 0.965 and for the test model it will be 0.881.

```{r}
summary_MLR <- rbind(summary_MLR3_train[-5], 
                     summary_MLR3_test[-5])
rownames(summary_MLR) <- c("Baseline MLR_Train", 
                           "Baseline MLR_Test" )
knitr::kable(summary_MLR, digits = 3)
```
Next we will be performing ridge and lasso regression model using the following code. We will split the data to train and test.
```{r}
# Training Data
data <- training
train.x <- data.matrix(data[,-ncol(data)])
train.y <- data.matrix(data[,ncol(data)])

# Test Data
data <- test
test.x <- data.matrix(data[,-ncol(data)])
test.y <- data.matrix(data[,ncol(data)])
```

Next we will perform the Ridge regression model.
```{r}
# Fit Ridge Regression Model
set.seed(123)
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, 
                    alpha = 0, lambda = cv_ridge$lambda.min)

# Evaluate Ridge Regression Model on Training Set
fit <- predict(glm_ridge, newx = train.x, type = "response")
true <- train.y
summary_ridge_train <- eval_results(true, fit)

# Evaluate Ridge Regression Model on Test Set
fit <- predict(glm_ridge, newx = test.x, type = "response")
true <- test.y
summary_ridge_test <- eval_results(true, fit)

# Summarise Ridge Regression Model Performance into Table
summary_ridge <- rbind(summary_ridge_train, summary_ridge_test)
rownames(summary_ridge) <- c("Ridge Model_Train", "Ridge Model_Test" )
```

Next we will perform Lasso Regression model

```{r}
# Fit LASSO Model
set.seed(123)
cv_lasso <- cv.glmnet(train.x,train.y, alpha = 1)
glm_lasso <- glmnet(train.x,train.y, alpha = 1, 
                    lambda = cv_lasso$lambda.min)

# Evaluate LASSO Model on Training Set
fit <- predict(glm_lasso, newx = train.x)
true <- train.y
summary_lasso_train <- eval_results(true, fit)

# Evaluate LASSO Model on Test Set
fit <- predict(glm_lasso, newx = test.x)
true <- test.y
summary_lasso_test <- eval_results(true, fit)

# Summarise LASSO Model Performance into Table
summary_lasso <- rbind(summary_lasso_train, summary_lasso_test)
rownames(summary_lasso) <- c("LASSO Model_Train", "LASSO Model_Test" )
```

We will compare the 3 models that we have build, Multi linear ,Ridge and Lasso regression models. We evaluate the 3 models together and compare both train and test models of each of the 3 regression model.

```{r}
summary_3models <- rbind(summary_MLR3_train, 
                         summary_ridge_train,                               
                         summary_lasso_train, 
                         summary_MLR3_test,                             
                         summary_ridge_test, 
                         summary_lasso_test)

rownames(summary_3models) <- c("Baseline MLR_Train", "Ridge Model_Train", 
                               "LASSO Model_Train",  "Baseline MLR_Test",  
                               "Ridge Model_Test", "LASSO Model_Test" )

knitr::kable(summary_3models[1:3,-5], digits = 3)
knitr::kable(summary_3models[4:6,-5], digits = 3)

```
From the 3 models we can see that the original multi linear regression model is the best one with the highest R^squared value. However the Lasso Regression Model has the lowest MSE, MAE and RMSE. We can conclude that the reason for the first model having the highest R^squared value might be due to overfitting as many values will get removed. Next we will build other models to see if they work better.

### 3.2.2. Selectively remove variables with high missing rate (To be done)
#### 3.2.2.1. Model Fitting
#### 3.2.2.2. Assessment
#### 3.2.2.3. Interpretation

### 3.2.3. Update the data set as we select variables (To be done)
#### 3.2.3.1. Model Fitting
#### 3.2.3.2. Assessment
#### 3.2.3.3. Interpretation

### 3.2.4. Imputation
**1. What is imputation**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.

We use **Multivariate Imputation By Chained Equations (MICE)** as our primary tool for this technique, because of its wide acceptance in scientific studies (@imputWorks). In order to use this technique, we have to assume that the "missingness" of a field can be explained by the values in other columns, (e.g., If the countries is in North America, it's more likely to be missing as we have discovered in 2.2). The general ideas of the algorithm is iteratively predict the missing values based on other variables, and improve it after each iterations until predicted value converse to a stable point. The algorithmic details of MICE is very concisely explained in @micealgo.

Our parameter for this algorithm is fairly standard:

@@

#### 3.2.4.1. Model Fitting
##### A. Ordinary Linear Regression
We generated 5 sets of imputated (training) data. We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). With all the generated data sets.

```{r}
formula.str <- 'pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'

# Using only the first imputed data set
fit1 <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
summary(fit1)
fit1.pred <- predict(fit1, countries.test.4)

# Build models with all generated data set and pool the estimates
fit2 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
fit2.combined <- pool(fit2)
summary(fit2.combined)
# dummy lm model
fit2.dummy <- lm(as.formula(formula.str), data = countries1)
# replace coefficients of dummy model & predict
fit2.dummy$coefficients <- fit2.combined$pooled$estimate
fit2.pred <- predict(fit2.dummy, countries.test.4)
```
Helper function to calculate R-Squared
```{r}
r2 <- function(pred, orig) {
   RSS <- sum((pred - orig)^2)
   TSS <- sum((pred - mean(orig))^2)
   R2 <- 1 - RSS/TSS
   return (R2)
}

adjR2 <- function(pred, orig, k) {
  R2 <- r2(pred, orig)
  n <- length(pred)
  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
  return (adjr2)
}
```
Adjusted R-Squared on train and test
```{r}
k <- length(fit1$coefficients) - 1
fit1.train.r2 <- summary(fit1)$adj.r.squared
fit1.test.r2 <- adjR2(fit1.pred, countries.test.4$pov, k)
fit2.train.r2 <- pool.r.squared(fit2, adjusted = T)[,'est']
fit2.test.r2 <- adjR2(fit2.pred, countries.test.4$pov, k)
res <- data.frame(model = c("Single", "Pooled"), train = c(fit1.train.r2, fit2.train.r2), test = c(fit1.test.r2, fit2.test.r2))
kable(res)
```

The models seem to be over-fitting. We should perform model simplification.

##### B. Regularization

Prepare data
